---
title: "Rossmann Store Sales - Data Modeling (Random Forest)"
author: "Nhi Huynh & Ngoc Ngo"
date: "8/24/2020"
output: html_document
---

INTRODUCTION

Rossmann is a Germany drug store chain with over 3790 stores in Europe. In this problem set obtained from Kaggle.com, participants are given records of sales of each store on different days, from 01/01/2013 to 31/07/2015. Other business factors such as holidays, promotions and competitions. Our goal is to forecast the sales of the store 6 weeks in advance, from 08/01/2015 to 09/17/2015 based on the data fields provided. We apply the Random Forest method to build a model that predict sales for these 6 weeks. Our work is based on Darius Barušauskas's script: https://www.kaggle.com/raddar/random-forest-example.

```{r}
library(readr)
library(randomForest)
```

```{r}
set.seed(678)
```

Read the train, test and store data
```{r}
train <- read.csv("train.csv")
test  <- read.csv("test.csv")
store <- read.csv("store.csv")
```

Merge train and test data with store
```{r}
train_store <- merge(train,store)
test_store <- merge(test,store)
```

Change all NAs to zeros
```{r}
train_store[is.na(train_store)] <- 0
test_store[is.na(test_store)] <- 0
```

Look at train data's column names and details
```{r}
names(train_store)
str(train_store)
summary(train_store)
```

Look at test data's column names and details
```{r}
names(test_store)
str(test_store)
summary(test_store)
```

Look at stores that are opened in train data only
```{r}
train_store <- train_store[ which(train_store$Open=='1'),]
```

Seperate out the elements of the date column for the train set
```{r}
train_store$Date <- as.Date(train_store$Date,format = "%m/%d/%y")
train_store$month <- as.integer(format(train_store$Date, "%m"))
train_store$year <- as.integer(format(train_store$Date, "%y"))
train_store$day <- as.integer(format(train_store$Date, "%d"))
```

Remove the date column (since elements are extracted) 
```{r}
train_store <- train_store[,-3]
```

Seperate out the elements of the date column for the test set
```{r}
test_store$Date <- as.Date(test_store$Date,format = "%m/%d/%y")
test_store$month <- as.integer(format(test_store$Date, "%m"))
test_store$year <- as.integer(format(test_store$Date, "%y"))
test_store$day <- as.integer(format(test_store$Date, "%d"))
```

Remove the date column (since elements are extracted)
```{r}
test_store <- test_store[,-4]
```

Now we finished data cleaning before applying the Random Forest model to this dataset. However, we will first test it with the train data first. We will split the train data so that 70% is used for modeling and 30% is used for testing:
```{r}
index = sample(2,nrow(train_store),replace = TRUE, prob=c(0.7,0.3))
train70 <- train_store[index==1,]
train30 <- train_store[index==2,]
```

Only select the columns in train data that are strong predictor values for sales
```{r}
feature.names <- names(train70)[c(1,2,5:20)]
feature.names
```
Since the "Customer" variable is not included in the test dataset, we will also not include it in the model.

```{r}
for (f in feature.names) {
  if (class(train70[[f]])=="character") {
    levels <- unique(c(train70[[f]], train30[[f]]))
    train70[[f]] <- as.integer(factor(train70[[f]], levels=levels))
    train30[[f]]  <- as.integer(factor(train30[[f]],  levels=levels))
  }
}
```

Apply the Random Forest model
```{r}
model_demo <- randomForest(train70[,feature.names], 
          log(train70$Sales+1), # +1 is added to avoid undefined errors if value is 0
          ntree=17,
          sampsize=100000, 
          do.trace=TRUE)
```
The mean squared error is low, indicating that the model might have a high accuracy.

Predict sales and test the accuracy of the model on the train dataset
```{r}
pred_demo <- exp(predict(model_demo, train30)) -1
1 - mean(abs(train30$Sales - pred_demo)/pred_demo)
```

The accuracy of this model is 88.47%.
Thus, we apply this model to predict the sales for the test data.

```{r}
feature.names <- names(train_store)[c(1,2,5:20)]
feature.names
```

```{r}
for (f in feature.names) {
  if (class(train_store[[f]])=="character") {
    levels <- unique(c(train_store[[f]], test_store[[f]]))
    train_store[[f]] <- as.integer(factor(train_store[[f]], levels=levels))
    test_store[[f]]  <- as.integer(factor(test_store[[f]],  levels=levels))
  }
}
```

Apply the Random Forest model
```{r}
model <- randomForest(train_store[,feature.names], 
                    log(train_store$Sales+1), 
                    ntree=17,
                    sampsize=100000, 
                    do.trace=TRUE)
```

Predict Sales and save the results in a .csv document
```{r}
pred <- exp(predict(model,test_store)) -1
sales_forecast <- data.frame(Id=test_store$Id, Sales=pred)
write_csv(sales_forecast,"SalesForecastRossman.csv") 
```
